---
title: "Airbnb - TMR report"
subtitle: "Data analysis"
# author: "Radoslaw Panczak & Thomas Sigler"
author:
- Radoslaw Panczak & Thomas Sigler
- Queensland Centre for Population Research
date: "`r format(Sys.time(), '%d %B, %Y')`"
geometry: "left=2cm,right=2cm,top=2cm,bottom=2cm"
output:
  pdf_document: 
    fig_caption: yes
    highlight: haddock
    keep_tex: yes
    number_sections: yes
    toc: yes
    toc_depth: 2
header-includes: 
- \usepackage{graphicx}
- \usepackage{float}
- \usepackage{pdflscape}
---

<!-- ------------------------------------------------------------ --> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      cache = FALSE,
                      fig.align="center",
                      fig.pos="H",
                      dpi = 300)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())

set.seed(12345)
options(scipen = 999)

library(pacman) 
p_load(tidyverse, magrittr, lubridate, scales, janitor, imputeTS,
       sf, tmap, tmaptools, 
       sjPlot, sjmisc, kableExtra, gghighlight)

tmap_mode("plot") 
```

<!-- ------------------------------------------------------------ --> 

```{r include=FALSE}
SA2 <- readRDS("TMR/data/SA2.Rds")
SA2_centr <- readRDS("TMR/data/SA2_centr.Rds")
STE <- readRDS("TMR/data/STE.Rds")
monthly_new_prep_tmr <- readRDS("TMR/data/monthly_new_prep_tmr.Rds")
```

<!-- ------------------------------------------------------------ --> 

\newpage

# Active Airbnb listings

State as of `2019-02-01`

```{r, include=FALSE}
temp <- monthly_new_prep_tmr %>% 
  group_by(SA2_MAIN16) %>% 
  summarize(mean_occupancy_rate = mean(occupancy_rate, na.rm = TRUE),
            median_occupancy_rate = median(occupancy_rate, na.rm = TRUE))

airbnb_sa2_occupancy_rate <- left_join(st_drop_geometry(SA2), temp) %>% 
  expand(SA2_MAIN16) %>% 
  filter(!is.na(SA2_MAIN16)) %>% 
  left_join(st_drop_geometry(SA2)) %>% 
  left_join(temp) %>% 
  mutate_if(is.factor, as.character) %>% 
  ungroup() %>% 
  arrange(SA2_MAIN16)

n <- monthly_new_prep_tmr %>% 
  filter(reporting_month == ymd("2019-02-01")) %>% 
  group_by(SA2_MAIN16) %>% 
  summarize(n = n())

airbnb_sa2_occupancy_rate %<>% 
  left_join(n) %>% 
  tidyr::replace_na(list(n = 0)) 

# summary(airbnb_sa2_occupancy_rate$mean_occupancy_rate)
# summary(airbnb_sa2_occupancy_rate$median_occupancy_rate)
stopifnot(nrow(SA2) == nrow(airbnb_sa2_occupancy_rate))

rm(temp, n)
```

```{r, fig.height = 3 , fig.width = 4}

descr(airbnb_sa2_occupancy_rate$n)

airbnb_sa2_occupancy_rate %>%
  dplyr::select(n) %>%
  plot_frq(type = "hist", show.mean = TRUE, geom.size = 10, show.sd = TRUE,
           ylim = c(0, 200)) +
  xlab("Number of properties in SA2") +
  ylab("Number of SA2 areas")

```

\newpage

```{r, fig.height = 8 , fig.width = 6 }
SA2_centr %<>% 
  left_join(select(airbnb_sa2_occupancy_rate, SA2_MAIN16,
                   n, mean_occupancy_rate, median_occupancy_rate))
tm_shape(SA2) +
  tm_borders(lwd = 0.5) +
  tm_shape(STE) +
  tm_borders(lwd = 1.0) +
  tm_shape(SA2_centr) +
  tm_bubbles(size = "n", 
             col = "darkorchid4", border.lwd = 0.1, border.col = "white",
             legend.size.is.portrait = TRUE,
             title.size = "Number of listings in SA2. \nState on 2019-02-01"
  ) +
  tm_credits("Geodata (c) ABS; data (c) Airdna", position=c("RIGHT", "BOTTOM"), size = 0.5) +
  tm_layout("Number of Airbnb listings",
            # legend.title.size = 1,
            # legend.text.size = 0.6,
            legend.position = c("right","top"),
            # legend.bg.color = "white",
            legend.bg.alpha = 1)
```



\newpage

```{r, fig.height = 8 , fig.width = 6 }
SA2_centr_zoom <- SA2_centr %>% 
  filter(GCC_NAME16 %in% c("Greater Brisbane", "Rest of NSW") | SA4_NAME16 == "Sunshine Coast")

SA2_zoom <- SA2 %>% 
  filter(GCC_NAME16 %in% c("Greater Brisbane", "Rest of NSW") | SA4_NAME16 == "Sunshine Coast")

tm_shape(SA2_zoom) +
  tm_borders(lwd = 0.5) +
  tm_shape(STE) +
  tm_borders(lwd = 1.0) +
  tm_shape(SA2_centr_zoom) +
  tm_bubbles(size = "n", 
             col = "darkorchid4", border.lwd = 0.1, border.col = "white",
             legend.size.is.portrait = TRUE,
             title.size = "Number of properties in SA2. \nState on 2019-02-01") +
  tm_credits("Geodata (c) ABS; data (c) Airdna", position=c("RIGHT", "BOTTOM"), size = 0.5) +
  tm_layout("Number of Airbnb listings",
            # legend.title.size = 1,
            # legend.text.size = 0.6,
            legend.position = c("left","bottom"),
            legend.bg.color = "white",
            legend.bg.alpha = 1)
```

\begin{landscape}

```{r, fig.height = 6 , fig.width = 8 }
SA2_centr_zoom <- SA2_centr %>% 
  filter(SA4_NAME16 == "Brisbane Inner City")

SA2_zoom <- SA2 %>% 
  filter(SA4_NAME16 == "Brisbane Inner City")

tm_shape(SA2_zoom) +
  tm_borders(lwd = 0.5) +
  tm_shape(STE) +
  tm_borders(lwd = 1.0) +
  tm_shape(SA2_centr_zoom) +
  tm_bubbles(size = "n", 
             col = "darkorchid4", border.lwd = 0.1, border.col = "white",
             legend.size.is.portrait = TRUE,
             title.size = "Number of properties in SA2. \nState on 2019-02-01"
  ) +
  tm_credits("Geodata (c) ABS; data (c) Airdna", position=c("RIGHT", "BOTTOM"), size = 0.5) +
  tm_layout("Number of Airbnb listings",
            # legend.title.size = 1,
            # legend.text.size = 0.6,
            legend.position = c("left","top"),
            legend.bg.color = "white",
            legend.bg.alpha = 1)
```

\end{landscape}

<!-- ------------------------------------------------------------ --> 

\newpage

# Mean occupancy of Airbnb listings

Average of individual listings in an area; using data from `2016-01-01` to `2019-02-01` period.

```{r, fig.height = 3 , fig.width = 4}

descr(airbnb_sa2_occupancy_rate$mean_occupancy_rate)

airbnb_sa2_occupancy_rate %>%
  dplyr::select(mean_occupancy_rate) %>%
  plot_frq(type = "hist", show.mean = TRUE, geom.size = 0.02, show.sd = TRUE,
           ylim = c(0, 100)) +
  xlab("Mean occupancy rate in SA2") +
  ylab("Number of SA2 areas")

```

\newpage

```{r, fig.height = 8 , fig.width = 6 }
tm_shape(SA2) +
  tm_borders(lwd = 0.5) +
  tm_shape(STE) +
  tm_borders(lwd = 1.0) +
  tm_shape(SA2_centr) +
  tm_bubbles(size = "n", 
             col = "mean_occupancy_rate", 
             n = 5, palette = "RdYlGn",
             border.lwd = 0.1, border.col = "white",
             legend.size.is.portrait = TRUE,
             title.size = "Number of listings (2019-02-01)",
             title.col = "Mean occupancy of listings in SA2. \nAverage from 2016-01-01 to 2019-02-01") +
  tm_credits("Geodata (c) ABS; data (c) Airdna", position=c("RIGHT", "BOTTOM"), size = 0.5) +
  tm_layout("Occupancy of Airbnb listings",
            # legend.title.size = 1,
            # legend.text.size = 0.6,
            legend.position = c("right","top"),
            # legend.bg.color = "white",
            legend.bg.alpha = 1)
```

\newpage

```{r, fig.height = 8 , fig.width = 6 }
SA2_centr_zoom <- SA2_centr %>% 
  filter(GCC_NAME16 %in% c("Greater Brisbane", "Rest of NSW") | SA4_NAME16 == "Sunshine Coast")

SA2_zoom <- SA2 %>% 
  filter(GCC_NAME16 %in% c("Greater Brisbane", "Rest of NSW") | SA4_NAME16 == "Sunshine Coast")

tm_shape(SA2_zoom) +
  tm_borders(lwd = 0.5) +
  tm_shape(STE) +
  tm_borders(lwd = 1.0) +
  tm_shape(SA2_centr_zoom) +
  tm_bubbles(size = "n", 
             col = "mean_occupancy_rate", 
             n = 5, palette = "RdYlGn",
             border.lwd = 0.1, border.col = "white",
             legend.size.is.portrait = TRUE,
             title.size = "Number of listings (2019-02-01)",
             title.col = "Mean occupancy of listings in SA2. \nAverage from 2016-01-01 to 2019-02-01") +
  tm_credits("Geodata (c) ABS; data (c) Airdna", position=c("RIGHT", "BOTTOM"), size = 0.5) +
  tm_layout("Occupancy of Airbnb listings",
            # legend.title.size = 1,
            # legend.text.size = 0.6,
            legend.position = c("left","bottom"),
            # legend.bg.color = "white",
            legend.bg.alpha = 1)
```

\begin{landscape}

```{r, fig.height = 6 , fig.width = 8 }
SA2_centr_zoom <- SA2_centr %>% 
  filter(SA4_NAME16 == "Brisbane Inner City")

SA2_zoom <- SA2 %>% 
  filter(SA4_NAME16 == "Brisbane Inner City")

tm_shape(SA2_zoom) +
  tm_borders(lwd = 0.5) +
  tm_shape(STE) +
  tm_borders(lwd = 1.0) +
  tm_shape(SA2_centr_zoom) +
  tm_bubbles(size = "n", 
             col = "mean_occupancy_rate", 
             n = 5, palette = "RdYlGn",
             border.lwd = 0.1, border.col = "white",
             legend.size.is.portrait = TRUE,
             title.size = "Number of listings (2019-02-01)",
             title.col = "Mean occupancy of listings in SA2. \nAverage from 2016-01-01 to 2019-02-01") + 
  tm_credits("Geodata (c) ABS; data (c) Airdna", position=c("RIGHT", "BOTTOM"), size = 0.5) +
  tm_layout("Number of Airbnb listings",
            # legend.title.size = 1,
            # legend.text.size = 0.6,
            legend.position = c("left","top"),
            # legend.bg.color = "white",
            legend.bg.alpha = 1)
```

\end{landscape}

<!-- ------------------------------------------------------------ --> 

\newpage

\begin{landscape}

\section{Temporal trends}

\subsection{Number of Airbnb listings}

```{r, include=FALSE}
airbnb_sa2_month_agg <- monthly_new_prep_tmr %>% 
  group_by(reporting_month) %>% 
  summarize(supply = n(),
            mean_occupancy_rate = mean(occupancy_rate, na.rm = TRUE))
```

```{r, fig.height = 6 , fig.width = 8 }
airbnb_sa2_month_agg %>%
  ggplot(aes(reporting_month, supply)) + 
  geom_col() +
  ylab("Number of listings in SA2") +
  xlab("")
```

\subsection{Mean occupancy rate}

```{r, fig.height = 6 , fig.width = 8 }
airbnb_sa2_month_agg %>%
  ggplot(aes(reporting_month, mean_occupancy_rate)) + 
  geom_col() +
  ylab("Mean occupancy rate in SA2") +
  xlab("")
```

\subsection{Distribution of occupancy rate}

```{r, fig.height = 6 , fig.width = 8 }
monthly_new_prep_tmr %>%
  ggplot(aes(reporting_month, occupancy_rate, group = reporting_month)) + 
  geom_boxplot() +
  ylab("Occupancy rate in SA2") +
  xlab("")
```

\subsection{Number of Airbnb listings by number of rooms}

```{r, include=FALSE}
airbnb_sa2_month_bed_agg <- monthly_new_prep_tmr %>% 
  group_by(reporting_month, bedrooms) %>% 
  summarize(supply = n())
```

```{r, fig.height = 6 , fig.width = 8 }
airbnb_sa2_month_bed_agg %>%
  ggplot(aes(reporting_month, supply, fill = as.factor(bedrooms))) + 
  geom_col(position = "dodge") +
  scale_fill_viridis_d(direction = -1) +
  ylab("Number of listings in SA2") +
  xlab("")  + labs(fill = "No of bedrooms")
```

\subsection{Number of Airbnb listings by listing type}

```{r, include=FALSE}
airbnb_sa2_month_list_agg <- monthly_new_prep_tmr %>% 
  group_by(reporting_month, listing_type) %>% 
  summarize(supply = n())
```

```{r, fig.height = 6 , fig.width = 8 }
airbnb_sa2_month_list_agg %>%
  ggplot(aes(reporting_month, supply, fill = as.factor(listing_type))) + 
  geom_col(position = "dodge") +
  scale_fill_viridis_d(direction = 1) +
  ylab("Number of listings in SA2") +
  xlab("")  + labs(fill = "Listing type")
```

\subsection{Number of Airbnb listings by listing type}

```{r, include=FALSE}
airbnb_sa2_month_prop_agg <- monthly_new_prep_tmr %>% 
  group_by(reporting_month, property_type_new) %>% 
  summarize(supply = n()) %>% 
  mutate(property_type_new = ifelse(property_type_new == "Unit", "Apartment", property_type_new)) %>% 
  mutate(property_type_new = ifelse(property_type_new == "", "Other", property_type_new))

```

```{r, fig.height = 6 , fig.width = 8 }
airbnb_sa2_month_prop_agg %>%
  ggplot(aes(reporting_month, supply, fill = as.factor(property_type_new))) + 
  geom_col(position = "dodge") +
  scale_fill_viridis_d(direction = 1) +
  ylab("Number of listings in SA2") +
  xlab("")  + labs(fill = "Property type")
```

\subsection{Number of Airbnb listings by SA4 area}

```{r, include=FALSE}
airbnb_sa2_month_sa4_agg <- monthly_new_prep_tmr %>% 
  group_by(reporting_month, SA4_NAME16) %>% 
  summarize(supply = n()) %>% 
  ungroup()
```

```{r, fig.height = 6 , fig.width = 8 }
airbnb_sa2_month_sa4_agg %>%
  ggplot(aes(reporting_month, supply, color = as.factor(SA4_NAME16))) + 
  geom_line() +
  scale_color_viridis_d(direction = 1) +
  ylab("Number of listings in SA2") +
  xlab("")  + labs(colour = "SA4 name") +
  gghighlight(max(supply) > 900, label_key = SA4_NAME16)
```

\subsection{Number of Airbnb listings by SA2 area}

```{r, include=FALSE}
airbnb_sa2_month_sa2_agg <- monthly_new_prep_tmr %>% 
  group_by(reporting_month, SA2_NAME16) %>% 
  summarize(supply = n()) %>% 
  ungroup()
```

```{r, fig.height = 6 , fig.width = 8 }
airbnb_sa2_month_sa2_agg %>%
  ggplot(aes(reporting_month, supply, color = as.factor(SA2_NAME16))) + 
  geom_line() +
  scale_color_viridis_d(direction = 1) +
  ylab("Number of listings in SA2") +
  xlab("")  + labs(colour = "SA4 name") +
  gghighlight(max(supply) > 600, label_key = SA2_NAME16)
```

\end{landscape}

```{r, include=FALSE}

```
