---
title: "Airbnb - ancillary data on SA2 level"
# subtitle: "Data preparation"
author: "Radoslaw Panczak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
---

<!-- ------------------------------------------------------------ --> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      cache = TRUE,
                      fig.width=8, fig.height=6, 
                      out.width="800px", out.height="600px", 
                      dpi=300)

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
set.seed(12345)
options(scipen = 999)

library(pacman) 
p_load(tidyverse, magrittr, lubridate, scales, readxl, janitor, imputeTS,
       sf, tmap, tmaptools, 
       ggridges, skimr, summarytools, sjmisc, kableExtra, naniar)

tmap_mode("view") # makes map interactive
```

<!-- ------------------------------------------------------------ --> 

# SA2 areas

```{r include=FALSE}
SA2_raw <- st_read("./data/geo/1270055001_sa2_2016_aust_shape/SA2_2016_AUST.shp", 
                   stringsAsFactors = FALSE)
```

There are `r comma(nrow(SA2_raw))` SA2 areas in Australia. We excluded areas with no physical location (including `r comma(nrow(filter(SA2_raw, grepl("Migratory - Offshore - Shipping", SA2_NAME16))))` `Migratory–Offshore–Shipping` and `r comma(nrow(filter(SA2_raw, grepl("No usual address", SA2_NAME16))))` `No usual address` codes for each State and Territory). And `r comma(nrow(filter(SA2_raw, SA2_5DIG16 %in% c("91001", "91002", "91004", "11161"))))` remote areas: 

```{r}
SA2_raw %>% 
  filter(SA2_5DIG16 %in% c("91001", "91002", "91004", "11161")) %>% 
  select(SA2_NAME16, STE_NAME16) %>% 
  st_drop_geometry()
```

```{r geodata_load, include=FALSE}
SA2 <- readRDS(file = "./data/geo/SA2_2016_AUST_clean.rds") %>% 
  # filter(SA4_CODE16 != "901") %>% 
  filter(! SA2_5DIG16 %in% c("91001", "91002", "91004", "11161")) %>% 
  select(-starts_with("SA3")) %>% 
  select(-starts_with("SA1")) %>% 
  select(-starts_with("GCC")) %>% 
  select(-AREASQKM16) %>% 
  mutate(STE_CODE16 = as.numeric(as.character(STE_CODE16))) %>% 
  mutate(SA2_MAIN16 = as.numeric(as.character(SA2_MAIN16))) %>% 
  mutate(SA2_5DIG16 = as.numeric(as.character(SA2_5DIG16))) %>% 
  mutate(SA4_CODE16 = as.numeric(as.character(SA4_CODE16))) %>% 
  mutate(SA4_GCC_CODE16 = case_when(
    SA4_CODE16 %in% c(402, 403, 404) ~ 401,
    SA4_CODE16 %in% c(302:305, 310, 311, 313, 314) ~ 301,
    SA4_CODE16 %in% c(207:214) ~ 206,
    SA4_CODE16 %in% c(503:507) ~ 502,
    SA4_CODE16 %in% c(115:128 ) ~ 102,
    TRUE ~ SA4_CODE16
  )
  ) %>% 
  mutate(SA4_GCC_NAME16 = case_when(
    SA4_GCC_CODE16 == 401 ~ "Greater Adelaide",
    SA4_GCC_CODE16 == 301 ~ "Greater Brisbane",
    SA4_GCC_CODE16 == 206 ~ "Greater Melbourne",
    SA4_GCC_CODE16 == 502 ~ "Greater Perth",
    SA4_GCC_CODE16 == 102 ~ "Greater Sydney",
    TRUE ~ SA4_NAME16
  )
  ) %>% 
  select(-SA4_CODE16, -SA4_NAME16)

stopifnot(nrow(SA2) == 
            nrow(SA2_raw) - 
            nrow(filter(SA2_raw, grepl("Migratory - Offshore - Shipping", SA2_NAME16))) -
            nrow(filter(SA2_raw, grepl("No usual address", SA2_NAME16))) -
            nrow(filter(SA2_raw, SA2_5DIG16 %in% c("91001", "91002", "91004", "11161")))
)

rm(SA2_raw)
gc()
```

That leaves the initial set of `r comma(nrow(SA2))` areas.

<!-- ------------------------------------------------------------ --> 

# Other explanatory variables

## Remotness

Data are from *1270.0.55.005 - Australian Statistical Geography Standard (ASGS): Volume 5 - Remoteness Structure, July 2016* [info](https://www.abs.gov.au/AUSSTATS/abs@.nsf/DetailsPage/1270.0.55.005July%202016?OpenDocument).

`No usual address` and `Migratory - Offshore - Shipping` categories are excluded. 

Data are on SA1 level. Population of SA1 was used to find majority classification within each SA2.

```{r include=FALSE}
remote <- read_csv("data/ABS/1270055005_ra_2016_aust_csv/RA_2016_AUST.zip", 
                   col_types = cols(AREA_ALBERS_SQKM = col_skip(), 
                                    MB_CODE_2016 = col_skip(), RA_CODE_2016 = col_integer(), 
                                    STATE_CODE_2016 = col_skip(), STATE_NAME_2016 = col_skip())) %>% 
  rename(SA1_MAIN16 = SA1_MAINCODE_2016,
         SA1_7DIG16 = SA1_7DIGITCODE_2016) %>% 
  group_by(SA1_MAIN16) %>% 
  filter(row_number()==1) %>% 
  ungroup() %>% 
  filter(!str_detect(RA_NAME_2016, "Migratory - Offshore - Shipping")) %>% 
  filter(!str_detect(RA_NAME_2016, "No usual address"))

pop_sa1 <- read_csv("data/ABS/2016 Census GCP Statistical Area 1 for AUST/2016Census_G01_AUS_SA1.csv.gz") %>% 
  select(SA1_7DIGITCODE_2016, Tot_P_P) %>% 
  rename(SA1_7DIG16 = SA1_7DIGITCODE_2016,
         pop_16 = Tot_P_P)

SA1 <- st_read("./data/geo/1270055001_sa1_2016_aust_shape/SA1_2016_AUST.shp", 
               stringsAsFactors = FALSE) %>% 
  st_drop_geometry() %>% 
  select(SA1_MAIN16, SA2_MAIN16, SA2_5DIG16, SA2_NAME16) %>% 
  mutate(SA1_MAIN16 = as.numeric(as.character(SA1_MAIN16))) %>% 
  mutate(SA2_MAIN16 = as.numeric(as.character(SA2_MAIN16))) %>% 
  mutate(SA2_5DIG16 = as.numeric(as.character(SA2_5DIG16)))

remote %<>% 
  left_join(pop_sa1) %>% 
  left_join(SA1)

remote %<>% 
  rename(pop_sa1_16 = pop_16) %>% 
  group_by(SA2_MAIN16) %>% 
  mutate(pop_sa2_16 = sum(pop_sa1_16)) %>% 
  ungroup() 
```

Original data

```{r}
remote %>% 
  frq(RA_NAME_2016)
```

SA2 assigned to majority

```{r}
remote_ra <- 
  remote %>% 
  group_by(SA2_MAIN16, RA_CODE_2016) %>% 
  mutate(pop_sa2_16_cat = sum(pop_sa1_16)) %>% 
  ungroup() %>% 
  mutate(pop_sa2_16_s = pop_sa2_16_cat/pop_sa2_16) %>% 
  arrange(SA2_MAIN16, desc(pop_sa2_16_s)) %>% 
  group_by(SA2_MAIN16, RA_CODE_2016) %>%
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  select(-SA1_MAIN16, -SA1_7DIG16, -pop_sa1_16) %>% 
  select(SA2_MAIN16, SA2_5DIG16, SA2_NAME16, everything())

remote_ra_one <- 
  remote_ra %>% 
  group_by(SA2_MAIN16) %>%
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  filter(! SA2_5DIG16 %in% c("91001", "91002", "91004", "11161"))
```

Share of majority category

```{r}
remote_ra_one %>% 
  descr(pop_sa2_16_s)

ggplot(remote_ra_one, aes(pop_sa2_16_s)) + 
  geom_histogram() +
  ylab("Number of SA2s") + xlab("Share of majority category") + theme_minimal()
```

Map 

```{r}
SA2 %>% 
  left_join(remote_ra_one) %>% 
  tm_shape() + 
  tm_polygons(col = "RA_NAME_2016", alpha = .5, border.col = NA, id = "SA3_NAME16")
```


<!-- ------------------------------------------------------------ --> 

## Level of urbanization

Data are from *1270.0.55.001 - Australian Statistical Geography Standard (ASGS): Volume 1 - Main Structure and Greater Capital City Statistical Areas, July 2016* [info](https://www.abs.gov.au/AUSSTATS/abs@.nsf/Lookup/1270.0.55.001Main+Features10018July%202016?OpenDocument)

`No usual address` and `Migratory - Offshore - Shipping` categories are excluded. 

Data are on SA1 level. Population of SA1 was used to find majority classification within each SA3.


```{r include=FALSE}
urban <- read_csv("data/ABS/1270055004_sa1_ucl_sosr_sos_2016_aust_csv/raw/SA1_UCL_SOSR_SOS_2016_AUST.zip", 
                  col_types = cols(AREA_ALBERS_SQKM = col_skip(), 
                                   SOSR_CODE_2016 = col_integer(), 
                                   SOS_CODE_2016 = col_integer(), 
                                   STE_CODE_2016 = col_skip(), STE_NAME_2016 = col_skip())) %>% 
  rename(SA1_MAIN16 = SA1_MAINCODE_2016,
         SA1_7DIG16 = SA1_7DIGITCODE_2016) %>% 
  select(-starts_with("UCL")) %>% 
  filter(!str_detect(SOSR_NAME_2016, "Migratory - Offshore - Shipping")) %>% 
  filter(!str_detect(SOSR_NAME_2016, "No usual address"))

urban %<>% 
  left_join(pop_sa1) %>% 
  left_join(SA1)

urban %<>% 
  rename(pop_sa1_16 = pop_16) %>% 
  group_by(SA2_MAIN16) %>% 
  mutate(pop_sa2_16 = sum(pop_sa1_16)) %>% 
  ungroup() 

rm(pop_sa1, SA1)
gc()
```

### Section of State Range (SOSR) 

Original data

```{r}
urban %>% 
  frq(SOSR_NAME_2016)
```

SA2 assigned to majority

```{r}
urban_sosr <- 
  urban %>% 
  select(-SOS_CODE_2016, -SOS_NAME_2016) %>% 
  group_by(SA2_MAIN16, SOSR_CODE_2016) %>% 
  mutate(pop_sa2_16_cat = sum(pop_sa1_16)) %>% 
  ungroup() %>% 
  mutate(pop_sa2_16_s = pop_sa2_16_cat/pop_sa2_16) %>% 
  arrange(SA2_MAIN16, desc(pop_sa2_16_s)) %>% 
  group_by(SA2_MAIN16, SOSR_CODE_2016) %>%
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  select(-SA1_MAIN16, -SA1_7DIG16, -pop_sa1_16) %>% 
  select(SA2_MAIN16, SA2_5DIG16, SA2_NAME16, everything())

urban_sosr_one <- 
  urban_sosr %>% 
  group_by(SA2_MAIN16) %>%
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  filter(! SA2_5DIG16 %in% c(90101, 90102, 90104, 10803))
```

Share of majority category

```{r}
urban_sosr_one %>% 
  descr(pop_sa2_16_s)

ggplot(urban_sosr_one, aes(pop_sa2_16_s)) + 
  geom_histogram() +
  ylab("Number of SA3") + xlab("Share of majority category") + theme_minimal()
```

Map 

```{r}
SA2 %>% 
  left_join(urban_sosr_one) %>% 
  tm_shape() + 
  tm_polygons(col = "SOSR_NAME_2016", alpha = .5, border.col = NA, id = "SA3_NAME16")
```

### Section of State (SOS)

Original data

```{r}
urban %>% 
  frq(SOS_NAME_2016)
```

SA2 assigned to majority

```{r}
urban_sos <- 
  urban %>% 
  select(-SOSR_CODE_2016, -SOSR_NAME_2016) %>% 
  group_by(SA2_MAIN16, SOS_CODE_2016) %>% 
  mutate(pop_sa2_16_cat = sum(pop_sa1_16)) %>% 
  ungroup() %>% 
  mutate(pop_sa2_16_s = pop_sa2_16_cat/pop_sa2_16) %>% 
  arrange(SA2_MAIN16, desc(pop_sa2_16_s)) %>% 
  group_by(SA2_MAIN16, SOS_CODE_2016) %>%
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  select(-SA1_MAIN16, -SA1_7DIG16, -pop_sa1_16) %>% 
  select(SA2_MAIN16, SA2_5DIG16, SA2_NAME16, everything())

urban_sos_one <- 
  urban_sos %>% 
  group_by(SA2_MAIN16) %>%
  filter(row_number() == 1) %>% 
  ungroup() %>% 
  filter(! SA2_5DIG16 %in% c(90101, 90102, 90104, 10803))
```

Share of majority category

```{r}
urban_sos_one %>% 
  descr(pop_sa2_16_s)

ggplot(urban_sos_one, aes(pop_sa2_16_s)) + 
  geom_histogram() +
  ylab("Number of SA2s") + xlab("Share of majority category") + theme_minimal()
```

Map 

```{r}
SA2 %>% 
  left_join(urban_sos_one) %>% 
  tm_shape() + 
  tm_polygons(col = "SOS_NAME_2016", alpha = .5, border.col = NA, id = "SA3_NAME16")
```

<!-- ------------------------------------------------------------ --> 

## ABS Census 2016 data 

### HOSD Housing Suitability

*This variable is new to 2016 and is a measure of housing utilisation based on a comparison of the number of bedrooms in a dwelling with a series of household demographics, such as the number of usual residents, their relationship to each other, age and sex. The criteria are based on the Canadian National Occupancy Standard (CNOS). This variable can be used to identify if a dwelling is either under or over utilised.* 

(Info)[https://www.abs.gov.au/ausstats/abs@.nsf/Lookup/by%20Subject/2900.0~2016~Main%20Features~HOSD%20Household%20Suitability~10123]

```{r}
HOSD <- read_xlsx("data/ABS/misc/raw/SA2 by HOSD Housing Suitability.xlsx", skip = 8) %>% 
  clean_names() %>%
  remove_empty(c("rows", "cols")) %>% 
  dplyr::rename(SA2_NAME16 = `x2`) %>% 
  filter(!grepl("Migratory - Offshore - Shipping", SA2_NAME16)) %>% 
  filter(!grepl("Cells in this table", SA2_NAME16)) %>% 
  filter(! SA2_NAME16 %in% c("Christmas Island", "Cocos (Keeling) Islands", 
                             "Jervis Bay", "Norfolk Island", 
                             "SA2", "Total")) %>%  
  dplyr::mutate(four_needed = four_or_more_extra_bedrooms_needed / total,
                three_needed = three_extra_bedrooms_needed / total,
                two_needed = two_extra_bedrooms_needed / total,
                one_needed = one_extra_bedroom_needed / total,
                optimal = no_bedrooms_needed_or_spare / total,
                one_spare = one_bedroom_spare / total,
                two_spare = two_bedrooms_spare / total,
                three_spare = three_bedrooms_spare / total,
                four_spare = four_or_more_bedrooms_spare / total) %>% 
  dplyr::mutate(needed = (four_or_more_extra_bedrooms_needed  + 
                            three_extra_bedrooms_needed + 
                            two_extra_bedrooms_needed + 
                            one_extra_bedroom_needed) / total,
                spare = (one_bedroom_spare +
                           two_bedrooms_spare + 
                           three_bedrooms_spare + 
                           four_or_more_bedrooms_spare) / total)

# missing 
HOSD_sf <- SA2 %>% 
  left_join(HOSD)

# # checking for airbnbs
# airbnb_HOSD_mis <- monthly_new_prep_rent %>%
#   filter(! SA2_NAME16 %in% HOSD$SA2_NAME16) %>% 
#   distinct(property_id, SA2_NAME16) %>% 
#   group_by(SA2_NAME16) %>% 
#   summarise(count = n())
# rm(airbnb_HOSD_mis)

```

```{r}
HOSD_sf %>% 
  tm_shape() +
  tm_polygons(c("needed", "spare"), 
              n = 7, palette = "RdYlBu", id = "SA3_NAME16") +
  tm_facets(free.scales.fill = FALSE)
```

## SEIFA

```{r}
SEIFA <- read_xls("data/ABS/SEIFA/raw/2033055001 - sa2 indexes.xls", 
                  sheet = "Table 1", skip = 5, na = "-") %>% 
  clean_names() %>%
  remove_empty(c("rows", "cols")) %>% 
  dplyr::rename(SA2_MAIN16 = `x1`, SA2_NAME16 = `x2`, 
                IRSD_s = `score_3`, IRSD_d = `decile_4`,
                IRSAD_s = `score_5`, IRSAD_d = `decile_6`, 
                IER_s = `score_7`, IER_d = `decile_8`, 
                IEO_s = `score_9`, IEO_d = `decile_10`, 
                URP = `x11`) %>% 
  dplyr::mutate(IRSD_d = as.integer(IRSD_d),
                IRSAD_d = as.integer(IRSAD_d),
                IER_d = as.integer(IER_d),
                IEO_d = as.integer(IEO_d)) %>% 
  filter(! SA2_NAME16 %in% c("Christmas Island", "Cocos (Keeling) Islands", 
                             "Jervis Bay", "Norfolk Island")) # Other Teritories


# missing all seifa
SEIFA_mis_sf <- SA2 %>% 
  anti_join(select(SEIFA, SA2_MAIN16, URP))

# # checking for airbnbs
# airbnb_SEIFA_mis <- monthly_new_prep_rent %>%
#   filter(SA2_MAIN16 %in% unique(SEIFA_mis_sf$SA2_MAIN16))


# missing one of seifa
SEIFA_na <- anti_join(SEIFA, select(drop_na(SEIFA), SA2_MAIN16))

SEIFA_na_sf <- SA2 %>% 
  right_join(SEIFA_na)

```

### Areas with no SEIFA

`r nrow(SEIFA_mis_sf)` areas, usually non-residential or with no/low pop. 

```{r}
SEIFA_mis_sf %>% 
  select(-SA2_MAIN16, -SA2_5DIG16, -STE_CODE16, -SA4_GCC_CODE16) %>% 
  tm_shape() +
  tm_polygons(col = "red", alpha = 0.5, lwd = 1)

# SA2 %<>% 
#   filter(!SA2_MAIN16 %in% SEIFA_mis_sf$SA2_MAIN16)
```

```{r}
# # checking for airbnbs
# airbnb_SEIFA_na <- monthly_new_prep_rent %>%
#   filter(SA2_MAIN16 %in% unique(SEIFA_na$SA2_MAIN16))

# There are `r comma(nrow(airbnb_SEIFA_mis))` records of `r comma(length(unique(airbnb_SEIFA_mis$property_id)))` propoerties in these areas.

# There are `r number(nrow(airbnb_SEIFA_na))` records of `r number(length(unique(airbnb_SEIFA_na$property_id)))` propoerties in these areas. 
```

### Areas without one or more indices

`r nrow(SEIFA_na)` areas, usually with small pops. **Excluded from analyses!**.

```{r}
SEIFA_na_sf %>% 
  select(-SA2_MAIN16, -SA2_5DIG16, -STE_CODE16, -SA4_GCC_CODE16) %>% 
  tm_shape() +
  tm_polygons(col = "red", alpha = 0.5, lwd = 1)
# SA2 %<>% 
#   filter(!SA2_MAIN16 %in% SEIFA_na$SA2_MAIN16)
```

```{r include=FALSE}
rm(SEIFA_na, SEIFA_na_sf, SEIFA_mis_sf)#, airbnb_SEIFA_na, airbnb_SEIFA_mis)
gc()
```


### IRSD

Index of Relative Socio-economic Disadvantage deciles 

```{r}
SEIFA_sa2_sf <- left_join(SA2, SEIFA)
rm(SEIFA)

tm_shape(SEIFA_sa2_sf) +
  tm_polygons(col = "IRSD_d", n = 10, palette = "BrBG", alpha = 0.5, lwd = 0)
```

### IRSAD

Index of Relative Socio-economic Advantage and Disadvantage deciles

```{r}
tm_shape(SEIFA_sa2_sf) +
  tm_polygons(col = "IRSAD_d", n = 10, palette = "BrBG", alpha = 0.5, lwd = 0)
```

### IER

Index of Economic Resources deciles

```{r}
tm_shape(SEIFA_sa2_sf) +
  tm_polygons(col = "IER_d", n = 10, palette = "BrBG", alpha = 0.5, lwd = 0)
```

### IEO

Index of Education and Occupation deciles

```{r}
tm_shape(SEIFA_sa2_sf) +
  tm_polygons(col = "IEO_d", n = 10, palette = "BrBG", alpha = 0.5, lwd = 0)
```

## Distance to coast

### Binary

Binary indicator has been created for costal proximity. SA2s that touch the coast have been flagged. Note: this is very crude proxy. Doesn't work well for large areas and in the areas with river estuaries.  

```{r}
coast_bin_sa2 <- read_csv("./data/airdna/raw/Spatial_join/coast_bin_sa2.zip",
                          col_types = cols(COAST = col_integer())) %>% 
  select(SA2_MAIN16, COAST) %>% 
  rename(coast_bin = COAST)

SA2 <- left_join(SA2, coast_bin_sa2)
```

```{r}
coast_bin_sa2 %>% 
  frq(coast_bin)

rm(coast_bin_sa2)
```

```{r}
tm_shape(SA2) +
  tm_polygons(col = "coast_bin", n = 2, alpha = 0.5, lwd = 0.5)
```

### Average distance

Alternative solution is to calculate distance to coast for all properties and then average that over area. However that leaves the areas without properties without data.

```{r}
coast_near <- read_csv("./data/airdna/raw/Spatial_join/coast_near.zip") %>% 
  rename(property_id = Property_ID, coast_dist = NEAR_DIST) %>% 
  select(property_id, coast_dist)

coast_near %>% 
  select(coast_dist) %>% 
  summarytools::descr(stats = c("mean", "med", "min", "max")) 

SA1_2016_join_new <- readRDS("./data/airdna/clean/SA1_2016_join_new.Rds") %>% 
  dplyr::select(property_id, SA2_5DIG16)

coast_near <- left_join(coast_near, SA1_2016_join_new) %>% 
  group_by(SA2_5DIG16) %>% 
  summarise(cost_mean = mean(coast_dist),
            cost_median = median(coast_dist))

coast_near %>% 
  select(cost_mean, cost_median) %>% 
  summarytools::descr(stats = c("mean", "med", "min", "max")) 

SA2 <- left_join(SA2, coast_near)

rm(coast_near, SA1_2016_join_new)
```

```{r}
tm_shape(SA2) +
  tm_polygons(col = "cost_median", n = 5, alpha = 0.5, lwd = 0.5)
```

## Tourist accomodation

ABS dataset '8635.0 - Tourist Accommodation, Australia, 2015-16' [dataset](http://www.abs.gov.au/AUSSTATS/abs@.nsf/Lookup/8635.0Main+Features12015-16?OpenDocument) could be used to provide information on supply and deman of accomodation on SA2 level. Several indicators are available on **monthly basis for July 2015 - June 2016 period**:

Establishments 
Rooms 
Bed spaces 
Room nights occupied 
Room nights available 
Room occupancy rate
Guest nights occupied 
Guest nights available 
Bed occupancy rate
Takings from accommodation 
Average takings per room night occupied
Average takings per room night available

See external description for an example of 'Guest Nights Occupied' calculations.

# Session info

```{r}
sessionInfo()
```

