---
title: "Airbnb"
subtitle: "GNO on SA3 level"
author: "Radoslaw Panczak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
    toc_float: yes
---

<!-- ------------------------------------------------------------ --> 
<!-- ------------------------------------------------------------ --> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, 
                      fig.width=8, fig.height=6, dpi=300, 
                      out.width="800px", out.height="600px")

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
set.seed(12345)
options(scipen = 999)

library(pacman) 
p_load(tidyverse, magrittr, readxl, janitor, anytime, scales,
       sf, tmap, tmaptools, geofacet, ggrepel,
       skimr, summarytools, sjmisc, kableExtra, naniar,
       googlesheets)

tmap_mode("view") # makes map interactive
```

```{r geodata_load, include=FALSE}
# SA3
SA3 <- readRDS(file = "./data/geo/SA3_2016_AUST_clean.rds") %>% 
  mutate(SA3_CODE16 = as.numeric(SA3_CODE16), 
         STE_CODE16 = as.numeric(STE_CODE16)) %>% 
  select(-starts_with("SA4"), -starts_with("GCC"), -AREASQKM16)

SA3_centr <- readRDS(file = "./data/geo/SA3_centr.rds") %>% 
  mutate(SA3_CODE16 = as.numeric(SA3_CODE16))

# STATES
STE <- readRDS(file = "./data/geo/STE_2016_AUST_clean.rds") %>%  
  mutate(STE_CODE16 = as.numeric(as.character(STE_CODE16))) %>% 
  filter(STE_CODE16 != 9)

# qtm(STE)
```

<!-- ------------------------------------------------------------ --> 
<!-- ------------------------------------------------------------ --> 

# Data preparation

```{r, include=FALSE}
property_select <- readRDS("./data/airdna/clean/property_new_select.Rds") %>% 
  select(-starts_with("SA"), -starts_with("STE"), -latitude, -longitude)

SA1_2016_join_new <- readRDS("./data/airdna/clean/SA1_2016_join_new.rds") %>% 
  dplyr::select(property_id, starts_with("SA3"), starts_with("STE"))

monthly_select <- readRDS("./data/airdna/clean/monthly_new_select.Rds") %>% 
  filter(reporting_month >= anydate("2016-01-01")) %>% 
  left_join(SA1_2016_join_new)

rm(SA1_2016_join_new)
```

## Exclusions

Raw dataset consists of `r format(nrow(monthly_select), nsmall=0, big.mark=",")` monthly stats of `r format(length(unique(monthly_select$property_id)), nsmall=0, big.mark=",")` properties. **Only data from 2016-01-01 onwards are included!**

```{r, include=FALSE}
#### Other Territories
OT_monthly <- monthly_select %>% 
  dplyr::filter(STE_NAME16 == "Other Territories")

monthly_select %<>% 
  dplyr::filter(!property_id %in% unique(OT_monthly$property_id))

##### zero reservations
# sjmisc::descr(monthly_select, number_of_reservations)
# sjmisc::descr(monthly_select, reservation_days)

# temp <- monthly_select %>% 
#   dplyr::filter(number_of_reservations == 0 & reservation_days > 0)
# 
# temp <- monthly_select %>% 
#   dplyr::filter(number_of_reservations > 0 & reservation_days == 0)

zero_res <- monthly_select %>% 
  dplyr::filter(number_of_reservations == 0 & reservation_days == 0)

# sjmisc::descr(zero_res, revenue_usd)
# sjmisc::descr(zero_res, revenue_native)

zero_res %<>% 
  select(property_id, reporting_month)

monthly_select %<>% anti_join(zero_res)

#### zero reservations - technically still possible? 
zero_days <- monthly_select %>% 
  dplyr::filter(number_of_reservations == 0)

# sjmisc::descr(zero_days, reservation_days)

# missing max guests but info on bedrooms
max_guests_1 <- 
  left_join(monthly_select, property_select) %>% 
  filter(is.na(max_guests) & !is.na(bedrooms)) 

max_guests_2 <- 
  left_join(monthly_select, property_select) %>% 
  filter(is.na(max_guests) & is.na(bedrooms)) 
```

First, `r format(nrow(OT_monthly), nsmall=0, big.mark=",")` records 
of `r format(length(unique(OT_monthly$property_id)), nsmall=0, big.mark=",")` properties  located in Other Teritories were excluded. 

Then, `r format(nrow(zero_res), nsmall=0, big.mark=",")` records 
of `r format(length(unique(zero_res$property_id)), nsmall=0, big.mark=",")` properties 
with zero reservations (and zero reservation days!) were excluded from monthly data. 


Then, `r format(nrow(max_guests_1), nsmall=0, big.mark=",")` records 
of `r format(length(unique(max_guests_1$property_id)), nsmall=0, big.mark=",")` properties 
with missing info on `max_guests` had their `max_guests` racalculated to be 1*number of `bedrooms`.

For `r format(nrow(max_guests_2), nsmall=0, big.mark=",")` records 
of `r format(length(unique(max_guests_2$property_id)), nsmall=0, big.mark=",")` properties 
with missing info on `max_guests`  that also had `bedrooms` info missing, `max_guests` was racalculated to be `1`. 

Note there are 
`r format(nrow(zero_days), nsmall=0, big.mark=",")` monthly records with zero `number_of_reservations` but `reservation_days` > 0 - they stay inside dataset.


```{r, include=FALSE}
airbnb_new_bookings_monthly <- 
  left_join(monthly_select, property_select) %>% 
  select(-occupancy_rate, -available_days, -blocked_days, 
         -host_id, -created_date, -last_scraped_date,
         -bathrooms) %>% 
  mutate(max_guests = ifelse(is.na(max_guests) & !is.na(bedrooms), 
                             bedrooms, max_guests)) %>% 
  mutate(max_guests = ifelse(is.na(max_guests) & is.na(bedrooms), 
                             1, max_guests))  

# saveRDS(airbnb_new_bookings_monthly, file = "./data/airdna/clean/airbnb_new_bookings_monthly.rds")

rm(zero_res, OT_monthly, zero_days, monthly_select, property_select, max_guests_1, max_guests_2)
gc()
```

Final dataset of `r format(nrow(airbnb_new_bookings_monthly), nsmall=0, big.mark=",")`
monthly records of `r format(length(unique(airbnb_new_bookings_monthly$property_id)), nsmall=0, big.mark=",")` properties was aggregated to SA3 levels. 

Locations of Airbnb properties (points) were linked to ABS data of SA3 areas (polygons) from 2016. Spatial join in ArcGIS was used with `CLOSEST` option to capture locations that did not overlap with polygons (see example here  https://www.airbnb.com/rooms/19103554 - due to privacy reasons, locations are not exact). 

```{r}
SA3_na <- anti_join(SA3, select(airbnb_new_bookings_monthly, SA3_CODE16))
```

## Areas with no data

`r SA3_na$SA3_NAME16` area, has no Airbnbs. **Excluded from analyses!**.

```{r}
tm_shape(SA3_na) +
  tm_polygons(col = "red", alpha = 0.5, lwd = 1)

SA3 <- anti_join(SA3, st_drop_geometry(select(SA3_na, SA3_CODE16)))
SA3_centr <- anti_join(SA3_centr, st_drop_geometry(select(SA3_na, SA3_CODE16)))
```

```{r eval=FALSE, include=FALSE}
frq(airbnb_new_bookings_monthly, listing_type, sort.frq = "desc")
frq(airbnb_new_bookings_monthly, property_type, sort.frq = "desc")
```


## 'Filling' time series

Time series filled to inlcude 0s in months with no GNOs. 

```{r, include=FALSE}
temp <- airbnb_new_bookings_monthly %>% 
  mutate(temp = reservation_days * max_guests) %>% 
  group_by(SA3_CODE16, reporting_month) %>% 
  summarise(
    reservations = sum(number_of_reservations),
    days = sum(reservation_days),
    max_gno = sum(temp, na.rm = TRUE)
  )

airbnb_sa3 <- left_join(st_drop_geometry(SA3), temp) %>% 
  expand(SA3_CODE16, reporting_month) %>% 
  left_join(st_drop_geometry(SA3)) %>% 
  left_join(temp) %>% 
  tidyr::replace_na(list(reservations = 0)) %>% 
  tidyr::replace_na(list(days = 0)) %>% 
  tidyr::replace_na(list(max_gno = 0)) %>% 
  mutate_if(is.factor, as.character) %>% 
  group_by(SA3_CODE16) %>% arrange(reporting_month) %>% mutate(cumu_res = cumsum(reservations),
                                                               cumu_day = cumsum(days),
                                                               cumu_gno = cumsum(max_gno)) %>% 
  ungroup() %>% 
  arrange(SA3_CODE16, reporting_month) %>% 
  mutate(STE_SHORT = case_when(
    STE_NAME16 == "Australian Capital Territory" ~ "ACT",
    STE_NAME16 == "New South Wales" ~ "NSW",
    STE_NAME16 == "Northern Territory" ~ "NT",
    STE_NAME16 == "Queensland" ~ "QLD",
    STE_NAME16 == "South Australia" ~ "SA",
    STE_NAME16 == "Tasmania" ~ "TAS",
    STE_NAME16 == "Victoria" ~ "VIC",
    STE_NAME16 == "Western Australia" ~ "WA"
  )) %>% 
  select(reporting_month, SA3_CODE16, SA3_NAME16, STE_SHORT, everything(), -STE_NAME16, -STE_CODE16)

# summarytools::freq(airbnb_sa3$reporting_month)
stopifnot(nrow(SA3) * length(unique(airbnb_new_bookings_monthly$reporting_month)) == nrow(airbnb_sa3))
```

```{r include=FALSE}
rm(temp, SA3_na)
gc()
```

# Prepared dataset

## 2016+ monthly

```{r fig.width=9, fig.height=6, dpi=300, out.width="900px", out.height="600px"}
ggplot(data = airbnb_sa3, 
       aes(x = reporting_month, y = max_gno/1000, 
           color = STE_SHORT, group = SA3_CODE16)) +
  geom_line(alpha=0.5) +
  theme_minimal() + xlab("") + ylab("Airbnb GNO [*1,000]") +
  scale_x_date(minor_breaks = "2 months")
```

## 2018 by state

```{r}
airbnb_sa3_18 <- airbnb_sa3 %>% 
  dplyr::select(-starts_with("cumu")) %>% 
  filter(reporting_month >= anydate("2018-01-01")) %>% 
  filter(reporting_month <= anydate("2018-12-01")) %>% 
  group_by(SA3_CODE16) %>% arrange(reporting_month) %>% mutate(year_gno = sum(max_gno)) %>% 
  ungroup() %>% 
  mutate(share_gno = (max_gno / year_gno) *100) %>% 
  arrange(SA3_CODE16, reporting_month)
```

```{r fig.width=9, fig.height=6, dpi=300, out.width="900px", out.height="600px"}
ggplot(airbnb_sa3_18, 
       aes(x = reporting_month, y = max_gno/1000, group = SA3_CODE16)) +
  geom_line(colour = "darkorchid4", alpha=0.5) +
  theme_minimal() + xlab("") + ylab("Airbnb GNO [*1,000]") +
  facet_geo(~ STE_SHORT, grid = "aus_grid1", scales = "free_y") +
  scale_x_date(minor_breaks = "1 month", date_breaks = "3 months", date_labels = "%b")
```

## 2018 relation to days

```{r}
ggplot(airbnb_sa3_18, aes(y = max_gno/1000, x = days/1000)) + 
  geom_smooth(method = lm, se = FALSE) +
  geom_point(aes(colour = STE_SHORT), alpha = 0.5) + 
  geom_rug() +
  theme_minimal() + ylab("GNO") + xlab("Airbnb reservation days") +
  facet_wrap(~reporting_month) 
```

## 2018 relation to reservations

```{r}
ggplot(airbnb_sa3_18, aes(y = max_gno/1000, x = reservations/1000)) + 
  geom_smooth(method = lm, se = FALSE) +
  geom_point(aes(colour = STE_SHORT), alpha = 0.5) + 
  geom_rug() +
  theme_minimal() + ylab("GNO") + xlab("Airbnb reservations") +
  facet_wrap(~reporting_month)

airbnb_sa3_18 %<>% 
  select(-reservations, -days)
```

## Relative 2018

```{r}
ggplot(data = airbnb_sa3_18, 
       aes(x = reporting_month, y = share_gno, group = SA3_CODE16, colour = sqrt(year_gno))) +
  geom_line(alpha=0.5) +
  theme_minimal() + xlab("") + ylab("Airbnb % of yearly GNO") +
  # facet_geo(~ STE_SHORT, grid = "aus_grid1") +
  facet_grid(~ STE_SHORT) +
  scale_colour_distiller(palette = "Spectral") +
  scale_x_date(minor_breaks = "1 month", date_breaks = "3 months", date_labels = "%b")
```

```{r}
saveRDS(airbnb_sa3, file = "./data/airdna/clean/airbnb_sa3.rds")
saveRDS(airbnb_sa3_18, file = "./data/airdna/clean/airbnb_sa3_18.rds")

```

<!-- ------------------------------------------------------------ --> 
<!-- ------------------------------------------------------------ --> 

# Areas of interest

## Locations

```{r eval=FALSE, include=FALSE}
gs_auth(new_user = TRUE)

regions <- gs_read(gs_title("TEMPO - regions of focus")) %>% 
  clean_names() %>%
  remove_empty(c("rows", "cols"))

saveRDS(regions, file = "./data/airdna/clean/regions.rds")

```
```{r}
regions <- readRDS(file = "./data/airdna/clean/regions.rds") %>% 
  select(sa3, region_types) %>% 
  rename(SA3_NAME16 = sa3)

SA3 %>% 
  right_join(regions) %>% 
  tm_shape() +
  tm_polygons(col = "region_types", id = "SA3_NAME16", alpha = 0.5, lwd = 1)
```

## Absolute

```{r}
regions %<>% 
  left_join(airbnb_sa3_18)

regions$SA3_NAME16_2 <- reorder(as.factor(regions$SA3_NAME16), as.numeric(as.factor(regions$region_types)))

ggplot(data = regions, 
       aes(x = reporting_month, y = max_gno, group = SA3_CODE16, fill = region_types)) +
  geom_col() +
  theme_minimal() + xlab("") + ylab("Airbnb yearly GNO") +
  facet_wrap(~SA3_NAME16_2, ncol = 3) +
  scale_fill_brewer(palette = "Dark2")
```

## Relative

```{r}
ggplot(data = regions, 
       aes(x = reporting_month, y = share_gno, group = SA3_CODE16, fill = region_types)) +
  geom_col() +
  theme_minimal() + xlab("") + ylab("Airbnb % of yearly GNO") +
  facet_wrap(~SA3_NAME16_2, ncol = 3) +
  scale_fill_brewer(palette = "Dark2")

```

<!-- ------------------------------------------------------------ --> 
<!-- ------------------------------------------------------------ --> 

# Monthly distribution of max GNO

## Distributions

Calculating proportion of GNO from one region & one month treated as 1.

```{r}
airbnb_sa3_prop_18 <- airbnb_sa3_18 %>% 
  dplyr::select(-year_gno, -share_gno) %>% 
  group_by(reporting_month) %>% 
  mutate(month_gno = sum(max_gno, na.rm = TRUE)) %>% 
  ungroup() %>% 
  mutate(share_gno = max_gno / month_gno) %>% 
  arrange(reporting_month, SA3_CODE16)

saveRDS(airbnb_sa3_prop_18, file = "./data/airdna/clean/airbnb_sa3_prop_18.rds")

ggplot(data = airbnb_sa3_prop_18, 
       aes(x = share_gno)) +
  geom_density() +
  theme_minimal() + ylab("Number of SA3s") + xlab("% of GNO") +
  facet_wrap(~reporting_month, ncol = 3) 
```

## Jan 

```{r fig.width=9, fig.height=7, dpi=300, out.width="900px", out.height="700px"}
SA3 %>% 
  left_join(filter(airbnb_sa3_prop_18,
                   reporting_month == anydate("2018-01-01"))) %>% 
  tm_shape() +
  tm_polygons(col = "share_gno", n = 7, palette = "-RdYlBu", id = "SA3_NAME16")
```


## Aug

```{r fig.width=9, fig.height=7, dpi=300, out.width="900px", out.height="700px"}
SA3 %>% 
  left_join(filter(airbnb_sa3_prop_18,
                   reporting_month == anydate("2018-08-01"))) %>% 
  tm_shape() +
  tm_polygons(col = "share_gno", n = 7, palette = "-RdYlBu", id = "SA3_NAME16")
```

## Jan vs Aug

```{r}
airbnb_sa3_prop_18 %>% 
  filter(reporting_month == anydate("2018-01-01") | 
           reporting_month == anydate("2018-08-01")) %>% 
  dplyr::select(SA3_CODE16, reporting_month, share_gno) %>% 
  spread(reporting_month, share_gno) %>% 
  ggplot(aes(x = `2018-01-01`, y = `2018-08-01`)) +
  geom_line(aes(x = `2018-08-01`, y = `2018-08-01`), colour = "darkorchid4", size = 1) +
  geom_point() +
  coord_fixed(xlim = c(0, 0.07), ylim = c(0, 0.07)) +
  theme_minimal() + 
    xlab("Jan") + ylab("Aug") 
```

## Jan vs Aug

```{r fig.width=9, fig.height=7, dpi=300, out.width="900px", out.height="700px"}
SA3 %>% 
  left_join(filter(airbnb_sa3_prop_18,
                   reporting_month == anydate("2018-01-01") | 
                     reporting_month == anydate("2018-08-01"))) %>% 
  tm_shape() +
  tm_dots(col = "share_gno", size = "max_gno", 
          n = 7, palette = "-RdYlBu", id = "SA3_NAME16") +
  tm_facets(by = "reporting_month")
```

# Session info

```{r}
sessionInfo()
```

