---
title: "Airbnb - income EDA 04"
subtitle: "Monthly relative income on SA4-GCC level"
# author: "Radoslaw Panczak"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: 
  html_document: 
    highlight: pygments
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 4
---

<!-- ------------------------------------------------------------ --> 
<!-- ------------------------------------------------------------ --> 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, fig.width=6, fig.height=3.75, dpi=300, out.width="600px", out.height="375px")

knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
set.seed(12345)
options(scipen = 999)

library(pacman) 
p_load(tidyverse, magrittr, readxl, janitor, lubridate, scales, 
       sf, tmap, tmaptools, 
       skimr, summarytools, sjmisc, kableExtra, naniar)

tmap_mode("view") # makes map interactive

isUnique <- function(vector){
  return(!any(duplicated(vector)))
}
```


```{r eval=FALSE, include=FALSE}
p_load(rmapshaper)

SA4_GCC <- st_read("./data/geo/SA4_GCC/SA4_GCC_2016_AUST.shp", stringsAsFactors = FALSE) 

SA4_GCC_clean <- ms_simplify(SA4_GCC, keep = 0.05, weighting = 0.7) # default settings

nrow(SA4_GCC_clean)
qtm(SA4_GCC_clean)

# st_write(SA4_GCC_clean, "data/geo/SA4_GCC_clean.shp")
saveRDS(SA4_GCC_clean, file = "data/geo/SA4_GCC_clean.rds")
```

```{r geodata_load, include=FALSE}
# SA4_GCC  combo done in arcgis
SA4_GCC <- readRDS(file = "./data/geo/SA4_GCC_clean.rds") %>% 
  filter(SA4_CODE16 != "901") %>% 
  mutate(SA4_CODE16 = as.numeric(as.character(SA4_CODE16))) %>% 
  mutate(STE_CODE16 = as.numeric(as.character(STE_CODE16))) %>% 
  rename(SA4_GCC_CODE16 = SA4_CODE16,
         SA4_GCC_NAME16 = SA4_NAME16)

# qtm(SA4_GCC)

# centroids from arcgis - with option 'inside'
SA4_GCC_centr <- st_read("./data/geo/SA4_GCC/SA4_GCC_2016_AUST_centr_inside.shp", stringsAsFactors = FALSE) %>% 
  filter(SA4_CODE16 != "901") %>% 
  mutate(SA4_CODE16 = as.numeric(as.character(SA4_CODE16))) %>% 
  mutate(STE_CODE16 = as.numeric(as.character(STE_CODE16))) %>% 
  rename(SA4_GCC_CODE16 = SA4_CODE16,
         SA4_GCC_NAME16 = SA4_NAME16)

# qtm(SA4_GCC_centr)

# STATES
STE <- readRDS(file = "./data/geo/STE_2016_AUST_clean.rds") %>%  
  mutate(STE_CODE16 = as.numeric(as.character(STE_CODE16))) %>% 
  filter(STE_CODE16 != 9)

# qtm(STE)

```

<!-- ------------------------------------------------------------ --> 
<!-- ------------------------------------------------------------ --> 

# Data 

## Preps

```{r include=FALSE}
airbnb_new_monthly <- readRDS(file = "./data/airdna/clean/monthly_new.rds")

# xxx <- airbnb_new_monthly %>% 
#   filter(is.na(SA4_CODE16)) %>% 
#   select(property_id) %>% 
#   unique()

airbnb_new_monthly %<>% 
  select(-starts_with("SA1"), -starts_with("SA2")) %>% 
  filter(!is.na(SA4_CODE16)) %>% 
  mutate(SA4_GCC_CODE16 = case_when(
    SA4_CODE16 %in% c(402, 403, 404) ~ 401,
    SA4_CODE16 %in% c(302:305, 310, 311, 313, 314) ~ 301,
    SA4_CODE16 %in% c(207:214) ~ 206,
    SA4_CODE16 %in% c(503:507) ~ 502,
    SA4_CODE16 %in% c(115:128 ) ~ 102,
    TRUE ~ SA4_CODE16
    )
  ) %>% 
  mutate(SA4_GCC_NAME16 = case_when(
    SA4_GCC_CODE16 == 401 ~ "Greater Adelaide",
    SA4_GCC_CODE16 == 301 ~ "Greater Brisbane",
    SA4_GCC_CODE16 == 206 ~ "Greater Melbourne",
    SA4_GCC_CODE16 == 502 ~ "Greater Perth",
    SA4_GCC_CODE16 == 102 ~ "Greater Sydney",
    TRUE ~ SA4_NAME16
    )
  )

# length(unique(SA4_GCC$SA4_GCC_CODE16))
# length(unique(airbnb_new_monthly$SA4_GCC_CODE16))
# tab(airbnb_new_monthly, SA4_NAME16, SA4_GCC_NAME16)
```


Prepared dataset of `r format(nrow(airbnb_new_monthly), nsmall=0, big.mark=",")`
monthly records of `r format(length(unique(airbnb_new_monthly$property_id)), nsmall=0, big.mark=",")`
properties was aggregated to combined SA4 & GCC levels. All SA4s from five metropolitan areas were aggregated and remaining SA4s were used "as is".

There are `r format(nrow(SA4_GCC), nsmall=0, big.mark=",")` SA4_GCC areas in Austrlia.

Locations of Airbnb properties (points) were linked to ABS data of SA areas (polygons) from 2016. Spatial join in ArcGIS was used with `CLOSEST` option to capture locations that did not overlap with polygons (see example here  https://www.airbnb.com/rooms/19103554 - due to privacy reasons, locations are not exact). 


```{r, include=FALSE}
temp <- airbnb_new_monthly %>% 
  group_by(SA4_GCC_CODE16, reporting_month) %>% 
  summarise(revenue = sum(revenue_native))

airbnb_sa4 <- left_join(st_drop_geometry(SA4_GCC), temp) %>% 
  expand(SA4_GCC_CODE16, reporting_month) %>% 
  filter(!is.na(reporting_month)) %>% 
  left_join(st_drop_geometry(SA4_GCC)) %>% 
  left_join(temp) %>% 
  tidyr::replace_na(list(revenue = 0)) %>% 
  mutate(revenue = as.integer(round(revenue))) %>% 
  mutate_if(is.factor, as.character) %>% 
  # group_by(SA4_GCC_CODE16) %>% arrange(reporting_month) %>% mutate(cumulative = cumsum(revenue)) %>% 
  ungroup() %>% 
  arrange(SA4_GCC_CODE16, reporting_month) %>% 
  select(-starts_with("GCC"))

# summarytools::freq(airbnb_sa1$reporting_month)
stopifnot(nrow(SA4_GCC) * length(unique(airbnb_new_monthly$reporting_month)) == nrow(airbnb_sa4))
rm(temp, airbnb_new_monthly)
gc()
```


## Time series

All SA4_GCC areas had Airbnb inside, but few regions are missing monthly observations on revenue. 

```{r}
airbnb_sa4 %>% 
  filter(revenue == 0) %>% 
  arrange(SA4_GCC_NAME16, reporting_month) %>% 
  select(SA4_GCC_NAME16, reporting_month, revenue)
```

Time series was filled with zeros for such cases.

Monthly revenue with `r format(nrow(airbnb_sa4), nsmall=0, big.mark=",")` data points of cumulative Airbnb revenue for `r format(nrow(SA4_GCC), nsmall=0, big.mark=",")` areas:

```{r}
ggplot(data=airbnb_sa4, 
       aes(x=reporting_month, y=revenue/1000000, 
           color=STE_NAME16, group=SA4_GCC_CODE16)) +
  geom_line(alpha=0.65) +
  theme_minimal() + xlab("") + ylab("Airbnb revenue [*1,000,000]$")

```

Relative revenue was calculated for 2018 data:

```{r}
airbnb_sa4_12mo <- airbnb_sa4 %>% 
  filter(reporting_month >= ymd("2018-01-01")) %>% 
  filter(reporting_month <= ymd("2018-12-01")) %>% 
  group_by(SA4_GCC_CODE16) %>% 
  mutate(revenue_total = sum(revenue)) %>% 
  ungroup() %>% 
  mutate(revenue_rel = (revenue/revenue_total)*100) %>% 
  arrange(SA4_GCC_NAME16, reporting_month) 

saveRDS(airbnb_sa4_12mo, (file = "./data/airdna/clean/airbnb_sa4_12mo.rds"))

ggplot(airbnb_sa4_12mo, 
       aes(x=reporting_month, y=revenue_rel, 
           color=STE_NAME16, group=SA4_GCC_CODE16)) +
  geom_line(alpha=0.5) +
  theme_minimal() + xlab("") + ylab("Airbnb relative revenue {%}")

```

This is the view across regions:

```{r}
temp <- left_join(SA4_GCC, airbnb_sa4_12mo)

tmap_mode("plot") 

tm_shape(temp) +
  tm_polygons(col = "revenue_rel", n = 5, style = "quantile", palette = "YlGnBu") +
  tm_facets(by = "reporting_month") +
  tm_shape(STE) +
  tm_borders("white", lwd=2)

tmap_save(tmap_last(), "plot/airbnb_sa4_12mo_seasonal.png", width=1920, height=1080, asp=0)

tmap_mode("view") 

rm(temp)
```


<!-- ------------------------------------------------------------ --> 
<!-- ------------------------------------------------------------ --> 

# Clustering temporal trajectories

## Methods

See Genolini *et al.* (2015) at https://www.jstatsoft.org/article/view/v065i04 

```{r, include=FALSE}
p_load(kml)

# kml needs wide data
airbnb_sa4_12mo_wide <- airbnb_sa4_12mo %>% 
  select(SA4_GCC_CODE16, SA4_GCC_NAME16, reporting_month, revenue_rel) %>% 
  mutate(Month = lubridate::month(reporting_month)) %>% 
  select((-reporting_month)) %>% 
  spread(Month, revenue_rel) %>% 
  select(SA4_GCC_CODE16, SA4_GCC_NAME16, `11`, `12`, `1`, `2`, `3`, `4`, `5`, `6`, `7`, `8`, `9`, `10`)

cld_airbnb <- cld(airbnb_sa4_12mo_wide, timeInData = 3:14, idAll = airbnb_sa4_12mo_wide$SA4_GCC_CODE16)

# kml(cld_airbnb, nbRedraw = 2, toPlot = "both")

kml(cld_airbnb, nbClusters=2:10)

# airbnb_sa4_12mo_wide$cluster2 <- getClusters(cld_airbnb, 2)
# airbnb_sa4_12mo_wide$cluster3 <- getClusters(cld_airbnb, 3)
airbnb_sa4_12mo_wide$cluster4 <- getClusters(cld_airbnb, 4)
# airbnb_sa4_12mo_wide$cluster5 <- getClusters(cld_airbnb, 5)
airbnb_sa4_12mo_wide$cluster6 <- getClusters(cld_airbnb, 6)

airbnb_sa4_12mo %<>% left_join(select(airbnb_sa4_12mo_wide, SA4_GCC_CODE16, starts_with("cluster")))

SA4_GCC %<>% left_join(select(airbnb_sa4_12mo_wide, SA4_GCC_CODE16, starts_with("cluster")))

```

## Number of clusters

No clear evidence for optimal amount of clusters

```{r}
# choice(cld_airbnb)
plotAllCriterion(cld_airbnb)
```

## Different cluster solutions

### 4 clusters

```{r}
ggplot(airbnb_sa4_12mo, 
       aes(x=reporting_month, y=revenue_rel)) +
  geom_line(aes(color=cluster4, group=SA4_GCC_CODE16), alpha=0.35) +
  geom_smooth(aes(group=cluster4, color=cluster4), size = 1, se = FALSE) +
  theme_minimal() + xlab("") + ylab("Airbnb relative revenue {%}") +
  scale_colour_brewer(palette = "Dark2")

ggplot(airbnb_sa4_12mo, 
       aes(x=reporting_month, y=revenue/1000000)) +
  geom_col(aes(fill=cluster4)) +
  theme_minimal() + xlab("") + ylab("Airbnb absolute revenue [*1,000,000]$")+ 
  scale_fill_brewer(palette = "Dark2") 

tabyl(SA4_GCC, cluster4)

tm_shape(SA4_GCC) +
  tm_polygons("cluster4", border.col = "white", lwd = 0.5, 
              id = "SA4_GCC_NAME16", popup.vars = c("cluster4"), 
              palette = "Dark2") +
  tm_shape(STE) +
  tm_borders(lwd = 0.5, alpha = 0.5)  
```

```{r}
tmap_mode("plot") 

tm_shape(SA4_GCC) +
  tm_polygons("cluster4", border.col = "white", lwd = 0.5, 
              id = "SA4_GCC_NAME16", popup.vars = c("cluster4"), 
              palette = "Dark2") +
  tm_shape(STE) +
  tm_borders(lwd = 0.5, alpha = 0.5) 

tmap_save(tmap_last(), "plot/airbnb_sa4_4clusters.png", 
          width=1600, height=1000)

tmap_mode("view") 

rm(temp)
```

### 6 clusters

```{r}
ggplot(airbnb_sa4_12mo, 
       aes(x=reporting_month, y=revenue_rel)) +
  geom_line(aes(color=cluster6, group=SA4_GCC_CODE16), alpha=0.35) +
  geom_smooth(aes(group=cluster6, color=cluster6), size = 1, se = FALSE) +
  theme_minimal() + xlab("") + ylab("Airbnb relative revenue {%}") +
  scale_colour_brewer(palette = "Dark2")

ggplot(airbnb_sa4_12mo, 
       aes(x=reporting_month, y=revenue/1000000)) +
  geom_col(aes(fill=cluster6)) +
  theme_minimal() + xlab("") + ylab("Airbnb absolute revenue [*1,000,000]$")+ 
  scale_fill_brewer(palette = "Dark2") 

tabyl(SA4_GCC, cluster6)

tm_shape(SA4_GCC) +
  tm_polygons("cluster6", border.col = "white", lwd = 0.5, 
              id = "SA4_GCC_NAME16", popup.vars = c("cluster6"), 
              palette = "Dark2") +
  tm_shape(STE) +
  tm_borders(lwd = 0.5, alpha = 0.5)  
```

# Geofacets of income

Using 4 clusters solution

## Absolute

```{r}
p_load(geofacet)

airbnb_sa4 %<>% 
  left_join(select(airbnb_sa4_12mo_wide, SA4_GCC_CODE16, 
                   starts_with("cluster"))) %>% 
  mutate(State_short = case_when(
    STE_NAME16 == "Australian Capital Territory" ~ "ACT",
    STE_NAME16 == "New South Wales" ~ "NSW",
    STE_NAME16 == "Northern Territory" ~ "NT",
    STE_NAME16 == "Queensland" ~ "QLD",
    STE_NAME16 == "South Australia" ~ "SA",
    STE_NAME16 == "Tasmania" ~ "TAS",
    STE_NAME16 == "Victoria" ~ "VIC",
    STE_NAME16 == "Western Australia" ~ "WA") )

# tabyl(airbnb_sa4, STE_NAME16)
# tabyl(airbnb_sa4, State_short)
```

```{r}
airbnb_ste <- airbnb_sa4 %>% 
  filter(reporting_month >= ymd("2017-06-01")) %>% 
  filter(reporting_month <= ymd("2019-02-01")) %>% 
  group_by(State_short, reporting_month, cluster4) %>% 
  summarise(revenue = sum(revenue)) %>% 
  ungroup() %>% 
  group_by(State_short, cluster4) %>% arrange(reporting_month) %>% mutate(cumulative = cumsum(revenue)) %>% 
  ungroup() %>% 
  arrange(State_short, cluster4, reporting_month)

airbnb_ste %>% 
  ggplot(aes(reporting_month, revenue/1000000)) +
  scale_fill_brewer(palette = "Dark2") + 
  scale_x_date(date_labels = "%y/%m") + 
  geom_col(aes(fill = cluster4)) +
  facet_geo(~ State_short, grid = "aus_grid1", scales = "free_y") +
  geom_text(aes(label = State_short), size = 3, col = "grey40", x = -Inf, y = Inf, hjust = -0.5, vjust = 1.2) +
  labs(
    title = "Monthly Airbnb absolute revenue in Australia",
    # caption = "Data Source: Airdna",
    y = "Revenue [*1,000,000$]", x = "") +
  theme_light() +
  theme(strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none")

```

## Relative

```{r}
airbnb_ste <- airbnb_sa4 %>% 
  filter(reporting_month >= ymd("2018-01-01")) %>% 
  filter(reporting_month <= ymd("2018-12-01")) %>% 
  group_by(State_short, reporting_month, cluster4) %>% 
  summarise(revenue = sum(revenue)) %>% 
  ungroup() %>% 
  group_by(State_short) %>% 
  mutate(revenue_total = sum(as.numeric(revenue))) %>% 
  ungroup() %>% 
  mutate(revenue_rel = (revenue/revenue_total)*100) %>% 
  mutate(revenue_rel = ifelse(revenue==0 & revenue_total==0, 0, revenue_rel)) %>% 
  arrange(State_short, cluster4, reporting_month) 

airbnb_ste %>% 
  ggplot(aes(reporting_month, revenue_rel)) +
  scale_fill_brewer(palette = "Dark2") + 
  geom_col(aes(fill = cluster4)) +
  geom_text(aes(label = State_short), size = 3, col = "grey40", x = -Inf, y = Inf, hjust = -0.5, vjust = 1.2) +
  scale_x_date(date_labels = "%y/%m") + 
  facet_geo(~ State_short, grid = "aus_grid1") +
  labs(
    title = "Monthly relative Airbnb revenue in Australia",
    # caption = "Data Source: Airdna",
    y = "% of yearly revenue", x = "") +
  theme_light() +
  theme(strip.background = element_blank(),
        strip.text = element_blank(),
        legend.position = "none")
```

<!-- ------------------------------------------------------------ --> 
<!-- ------------------------------------------------------------ --> 

```{r include=FALSE}
file.remove("cld_airbnb.Rdata")
```

```{r}
# Session info
# sessionInfo()
```

