## property_new type

```{r, eval=FALSE}
property_new %<>%
  mutate(property_type_orig = property_type) %>% 
  mutate(property_type = ifelse(property_type == "Bed &amp; Breakfast", "Bed & Breakfast", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Heritage hotel (India)", "Heritage hotel", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Serviced apartment", "Apartment", property_type)) %>% 
  mutate(property_type = ifelse(grepl("Private room in", property_type), "Private room", property_type)) %>% 
  mutate(property_type = ifelse(grepl("Shared room in", property_type), "Shared room", property_type)) %>% 
  # mutate(Created_Date = as.Date(Created_Date)) %>% 
  # mutate(Last_Scraped_Date = as.Date(Last_Scraped_Date)) %>% 
  # mutate(Calendar_Last_Updated = as.Date(Calendar_Last_Updated)) %>% 
  mutate(property_type = str_replace_all(property_type, 'Entire ', '')) %>% 
  mutate(property_type = R.utils::capitalize(property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Bed & breakfast", "Bed & Breakfast", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Earth House", "Earth house", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Van", "Camper/RV", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Loft", "Apartment", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Condominium", "House", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Townhouse", "House", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Guesthouse", "House", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Holiday home", "House", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Villa", "House", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Vacation home", "House", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Hostel", "Hotels", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Heritage hotel", "Hotels", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Dorm", "Hotels", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Boutique hotel", "Hotels", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Bed & Breakfast", "Hotels", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Tent", "Tents", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Yurt", "Tents", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Tipi", "Tents", property_type)) %>% 
  mutate(property_type = ifelse(property_type == "Igloo", "Tents", property_type))
```

`property_new type` variable was simplified:

* `Bed & Breakfast` spelling & typo variations were cleaned
* `Serviced apartment` and `Loft` were merged to `Apartment`
* All the combinations of `Private room in ...` were replaced to `Private room` 
* Same for `Shared room in ...` 
* Places names labelled as `Entire ...` had this word removed (see below for entire issue) 
* `Van` was merged with `Camper/RV` category
* `Condominium`, `Townhouse`, `Guesthouse`, `Holiday home`, `Villa` and `Vacation home` were merged to form category **`House`**
* `Hostel`, `Heritage hotel`, `Dorm`, `Boutique hotel`, `Bed & Breakfast` were merged to form category **`Hotels`**
* `Tent`, `Yurt`, `Tipi`, `Igloo` were merged to form category **`Tents`**

TODO >> further simplify?
* Cottage?
* Chalet?
* Nature lodge?
* Casa particular?

Still quite few categories with small numbers: 

```{r, fig.width=8, fig.height=8, out.width="800px", out.height="800px"}
# frq(property_new, property_type)
ggplot(property_new, aes(property_type))+ 
  geom_bar() + coord_flip() + theme_minimal()
```


